---
title: "Questions 3 and 4"
author: "Christopher"
date: "10/18/2021"
output: html_document
---

## Using the BRFSS dataset located at "https://github.com/kijohnson/ADA-Fall-2021/blob/master/BRFSS2017_10percent_v.csv",  a) import the dataset and show your code,  write code to b) determine median heights for each sex (males and females) and, c) make a plot to examine median differences in height by sex. Report the median heights and upload a PDF or picture of your plot.

```{r}
#Question 3a) import the dataset and show your code

library(haven) #This program is for reading the file from the specified link from github
library("foreign")
library(readr) #for reading csv file (read_csv())
## importing data from a csv file to r
BRFSS2017 <- read_csv("https://github.com/kijohnson/ADA-Fall-2021/blob/master/BRFSS2017_10percent_v.csv?raw=true") 

#knowing the variable type for both sex and height
class(BRFSS2017$SEX)
class(BRFSS2017$ht_meters)

#install.packages("tidyverse")
library(tidyverse)
library(dplyr)

# conducting data management for the sex and height in meters
#Summarizing sex to assess the distribution and see if there are any unusual values in the data
summary(BRFSS2017$SEX)

#recode sex as a factor variable
BRFSS2017$SEX<-factor(BRFSS2017$SEX, 
                      levels = c(1, 2), 
                      labels = c("Male", "Female"))

#summary of the recoded sex variable
table(na.omit(BRFSS2017$SEX))

#Assessing height distribution.
summary(BRFSS2017$ht_meters)


#Question 3b) Determining the median for each of the sex groups
summary( BRFSS2017$ht_meters[BRFSS2017$SEX=="Male"], na.rm=TRUE )
median( BRFSS2017$ht_meters[BRFSS2017$SEX=="Male"], na.rm=TRUE )

summary( BRFSS2017$ht_meters[BRFSS2017$SEX=="Female"], na.rm=TRUE )
median( BRFSS2017$ht_meters[BRFSS2017$SEX=="Female"], na.rm=TRUE )


#Question 3c) Examining median differences in height by sex using the boxplot
ggplot(data = na.omit(BRFSS2017), aes(x = SEX, y = ht_meters)) +
    geom_boxplot() +
  labs(x = "SEX", y = "Height (meters)",  title = "Boxplot for Sex and Height")
```



#Question 4 
#a. Put the version of the code file that you wrote that includes annotation for what each step does in a Github repository (you can use the one previously created in class or for your homework or create a new one).

#b. Next add code to determine the mean height in females and males. Save this new version and resubmit the new version to your Github repository.

#c. Provide a link to your Github repository containing the code file for review.
```{r}
#Determining the mean values for both Male and Female participants using the mean function
mean( BRFSS2017$ht_meters[BRFSS2017$SEX=="Male"], na.rm=TRUE ) #1.781511
mean( BRFSS2017$ht_meters[BRFSS2017$SEX=="Female"], na.rm=TRUE ) #1.631513

```

